How to select predictive models for decision-making or causal inference?

### 2023-12-16 how_to_select_causal_models.tex
iter

Given complex,
potentially noisy, data, which model is to be most trusted to yield valid causal
estimates? As no single learner performs best on all data sets, there is a
pressing need for clear guidelines to select outcome models for causal
inference.

@@

We study \textit{model selection procedures} in practical settings:
\textit{finite samples} without the \textit{well-specification}
assumptions which are often made in the literature (Appendix
\ref{apd:prior_work}). Asymptotic causal-inference theory calls for complex
risks, but a practical question is whether model-selection procedures, that rely
on data split, can estimate these risks reliably enough. Indeed, they come with
more quantities to estimate, which may bring additional variance, leading to
worse model selection.

We first illustrate the problem of causal model selection and briefly review
prior art. Then we anchor causal model selection in the \emph{potential outcome}
framework and details the causal risks and model-selection procedure. We then
rewrite the so-called $R\text{-risk}$, as a reweighted version of mean squared
difference between the true and estimated individualized treatment effect.
Finally, we conduct a thorough empirical study comparing the different metrics
on a wide variety of data-generative processes based on simulations and real
data, going beyond prior work based on specific simulation settings
\citep{schuler_comparison_2018,alaa_validating_2019} (Appendix
\ref{apd:prior_work}).

>>>>

Given complex health data, which model is to be most trusted to yield 
valid causal estimates needed to motivate individual treatment decisions?
As no single learner performs best on all data sets, there is a
pressing need for clear guidelines to select outcome models for causal
inference.

@@

We study \textit{model selection procedures} in practical settings,
without theoretical assumptions often made in statistical literature such
as \textit{infinite} data or \textit{well-specified} models (Appendix
\ref{apd:prior_work}). Asymptotic causal-inference theory recommends complex
risks, but a practical question is whether model-selection procedures, that rely
on data split, can estimate these risks reliably enough. Indeed, these
risks come with
more quantities to estimate, which may bring additional variance, leading to
worse model selection.

We first illustrate the problem of causal model selection and briefly review
prior art. Then we anchor causal model selection in the \emph{potential outcome}
framework and details the causal risks and model-selection procedure. We then
rewrite the so-called $R\text{-risk}$ as a reweighted version of mean squared
difference between the true and estimated individualized treatment effect.
Finally, we conduct a thorough empirical study comparing the different metrics
on diverse datasets, using a family of simulations and real health
data, going beyond prior work limited to specific simulation settings
\citep{schuler_comparison_2018,alaa_validating_2019} (Appendix
\ref{apd:prior_work}).
